\textbf{Elimination Of Jump Statements}

### 2022-05-10 paper/main.tex
Update main.tex

\documentclass[two column]{article}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{graphicx}

@@

@@

\begin{document}

@@

\section{Introduction}
The problem is known both computer science and software engineering communities for a few decades already, while it was firstly formulated by Dijkstra in 1968: he affirmed that the use of \textbf{goto} has disastrous effect and proposed to remove it from all programming languages, Even though a few years later Knuth showed a manner for structuring programs with the \textbf{goto} statement and argued that it could be a powerful tool if it is well used, researchers got interested in finding a way for restructuring programs and eliminating \textbf{goto} statements.

Williams created a tool for restructuring Pascal code; Jonsson was one of the first who made a summary of so called \textbf{GOTO}-patches, which are design techniques for eliminating \textbf{goto} statements; Erosa suggested an algorithm fir automatic elimination of \textbf{goto} statements in C; Zegour explained how \textbf{goto} elimination may work in any language, as long as it has jump statements and control structures; Morris suggested to use regular expressions for replacing \textbf{goto} with IF-THEN-ELSE constructs; Ramshow proposed an algorithm for Pascal programs; Ganapathi introduced optimisation method for SHIM IR; Ceccato suggested how to eliminate \textbf{goto} during migration of legacy code to Java,

@@

The result of our research is an algorithm that allows converting a program in EO (Elementary Objects) language containing \textbf{GOTO} objects into a semantically similar one, but without using them. We propose an algorithm that replaces code fragments using the GOTO object with similar fragments, but without using it. We claim that two algorithms are semantically equivalent if with the same set of input arguments they return the same results.

\section{Methods}
In EOLANG, all jump statements are implemented as a single \textbf{GOTO}-object. The \textbf{GOTO}-object represents two types of jumps: \textit{forward jump} and \textit{backward jump}. Operators such as \textbf{break}, \textbf{return} and \textbf{goto forward} are implemented via \textit{forward jump}; \textbf{continue} and \textbf{goto backward} are implemented via \textit{backward jump}.

For each of them, a semantically identical replacement is required so that the algorithm returns the same output with the same input parameters.

\subsection{forward jump}
\textbf{\textit{Goto forward}} is implemented in the language as:
\begin{lstlisting}
+import org.eolang.gray.goto
+import org.eolang.io.stdout

@@

In a program \textit{forward jump} can be used within the body of \textbf{if}-statement. In this case program will look like this:

\includegraphics{forward1}

Let's look at the following code on EOLANG:
\begin{lstlisting}
[x] > f
 memory > r
 seq > @
  r.write 0
  goto
   [g]
    seq > @
     if.
      x.eq 0
      g.forward
      TRUE
     r.write (42.div x)
  r
\end{lstlisting}

In this example we can see that the use of "forward jump" is unnecessary because it is called when condition within \textbf{if}-statement is TRUE. By inverting the condition in \textbf{if}-statement, you can write the code block between the \textbf{goto} call and its label into a separate \textbf{if}-statement:

\includegraphics{forward2}

@@

\subsubsection{return}
As mentioned above, return statement is also implemented via \textit{goto forward} object. Consider the following code:
\begin{lstlisting}
[a b] > max
 goto > @
  [g]
   seq > @
    if.
     a.greater b
     g.forward a
     TRUE
    b
\end{lstlisting}
When using goto as \textit{return} statement, an additional attribute is added to the \textbf{goto}-object, similar to the return value of a procedure. In this case, it can easily be transformed into an \textbf{if}-statement as shown in the paragraph above:
\begin{lstlisting}
[a b] > max
 if.
  a.greater b
  a
  TRUE
 b
\end{lstlisting}

\subsubsection{break}
Break is a control-flow statement which could also be implemented using \textbf{goto forward}. Since the break statement can only be found inside the generator (in EOlang Generator is a routine that can be used to control iteration behaviour of a loop) Its replacement is slightly different from the one presented in the paragraph above. In this case, we need to invert the \textit{if}-condition and add it to the condition of the loop with the logical operator \textit{AND}. After that, we easily get rid of the \textit{if}-condition itself.

@@

\subsection{backward jumps}
Backward jump is usually used similarly to a while loop and can be easily replaced by it (considering programs that are compiled and executed). Substitution will look like \textbf{do-while} construction.

\includegraphics{backward1}

@@

In this case we could easily replace construction using \textbf{goto} with the \textbf{while}:

\includegraphics{backward2}

@@

\subsubsection{continue}
Continue is also a control-flow statement, which occurs in loops and is implemented through the EO object \textbf{GOTO}. In order to replace it, we need to invert the \textit{if}-condition in the loop and all the statements that occur after \textbf{goto} are put in the \textit{if} body with the inverted condition.

@@

Replace the GOTO object with the described method:
\begin{lstlisting}
[]
 memory > i
 seq > @
 while.
  i.less 10
  seq
   i.write (i.add 1)
   if.
    not. (eq. (i.mod 5 0))
    stdout "Not divisible by 5!\n"
    TRUE
\end{lstlisting}

>>>>

\documentclass[two column]{article}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{xcolor}

@@

\definecolor{codegray}{rgb}{0.5,0.5,0.5}

\lstdefinestyle{mystyle}{
  %%backgroundcolor=\color{backcolour}, commentstyle=\color{codegreen},
  %%keywordstyle=\color{magenta},
  numberstyle=\tiny\color{codegray},
  %%stringstyle=\color{codepurple},
  basicstyle=\ttfamily\footnotesize,
  breakatwhitespace=false,         
  breaklines=true,
  frame=leftline,
  captionpos=b,                    
  keepspaces=true,                 
  numbers=left,                    
  numbersep=5pt,                  
  showspaces=false,                
  showstringspaces=false,
  showtabs=false,                  
  tabsize=2
}

@@

\lstset{style=mystyle}
\begin{document}

@@

\section{Introduction}
The problem is known both computer science and software engineering communities for a few decades already, while it was firstly formulated by Dijkstra in 1968: he affirmed that the use of GOTO has disastrous effect and proposed to remove it from all programming languages, Even though a few years later Knuth showed a manner for structuring programs with the GOTO statement and argued that it could be a powerful tool if it is well used, researchers got interested in finding a way for restructuring programs and eliminating GOTO statements.

Williams created a tool for restructuring Pascal code; Jonsson was one of the first who made a summary of so called GOTO-patches, which are design techniques for eliminating GOTO statements; Erosa suggested an algorithm fir automatic elimination of GOTO statements in C; Zegour explained how GOTO elimination may work in any language, as long as it has jump statements and control structures; Morris suggested to use regular expressions for replacing GOTO with IF-THEN-ELSE constructs; Ramshow proposed an algorithm for Pascal programs; Ganapathi introduced optimisation method for SHIM IR; Ceccato suggested how to eliminate GOTO during migration of legacy code to Java,

@@

The result of our research is an algorithm that allows converting a program in EO (Elementary Objects) language containing GOTO objects into a semantically similar one, but without using them. We propose an algorithm that replaces code fragments using the GOTO object with similar fragments, but without using it. We claim that two algorithms are semantically equivalent if with the same set of input arguments they return the same results.

\section{Methods}
In EOLANG, all jump statements are implemented as a single \textbf{goto}-object. The \textbf{goto}-object represents two types of jumps: \textit{forward jump} and \textit{backward jump}. Operators such as \textbf{break}, \textbf{return} and \textbf{goto forward} are implemented via \textit{forward jump}; \textbf{continue} and \textbf{goto backward} are implemented via \textit{backward jump}.

We need to find a solution for both, how to replace them in different situations. For each of them, a semantically identical replacement is required so that the algorithm returns the same output with the same input parameters.

\subsection{forward jump}
\textbf{Goto forward} is implemented in EOLANG as:
\begin{lstlisting}
+import org.eolang.gray.goto
+import org.eolang.io.stdout

@@

In a program \textit{forward jump} can be used within the body of \textit{if-statement}. Structure of an algorithm can be represented by a scheme on Figure 1. Inside \textit{if-statement} are located some statements (stmt1), after that \textbf{goto-forward}, some other statements (stmt2) and the label of \textbf{goto}, where the program will continue to work if the condition in \textit{if-statement} was equal to \textit{TRUE}.

\begin{figure}[h]
    \centering
    \includegraphics{forward1}
    \caption{\textit{forward jump} scheme}
    \label{fig:my_label}
\end{figure}

Let's look at the following code:
\begin{lstlisting}
[x] > f
 memory > r
 seq > @
  r.write 0
  goto
   [g]
    seq > @
     if.
      x.eq 0
      g.forward
      TRUE
     r.write (42.div x)
  r
\end{lstlisting}

In this example we can see that the use of "forward jump" is unnecessary because it is called when condition within \textit{if-statement} is TRUE. By inverting the condition, you can write the code block between the \textbf{goto} call and its label into a separate \textit{if-statement}. This approach is illustrated in Figure 2.

\begin{figure}[h]
    \centering
    \includegraphics{forward2}
    \caption{\textit{forward jump} elimination}
    \label{fig:my_label}
\end{figure}

@@

\subsubsection{return}
As mentioned above, \textbf{return} statement is also implemented via \textit{goto forward} object. Consider the following code:
\begin{lstlisting}
[a b] > max
 goto > @
  [g]
   seq > @
    if.
     a.greater b
     g.forward a
     TRUE
    b
\end{lstlisting}
When using goto as \textit{return} statement, an additional attribute is added to the \textbf{goto}-object, similar to the return value of a procedure. In this case, it can easily be transformed into an \textbf{if}-statement as shown in the paragraph above:
\begin{lstlisting}
[a b] > max
 if.
  a.greater b
  a
  TRUE
 b
\end{lstlisting}

\subsubsection{break}
\textit{Break} is a control-flow statement which could also be implemented using \textbf{goto forward}. Since the \textit{break} statement can only be found inside the generator (in EOlang Generator is a routine that can be used to control iteration behaviour of a loop) Its replacement is slightly different from the one presented in the paragraph above. In this case, we need to invert the \textit{if}-condition and add it to the condition of the loop with the logical operator \textit{AND}. After that, we easily get rid of the \textit{if}-condition itself.

@@

Thus, at each iteration of the \textit{while} loop, the added condition will be checked, and as soon as it becomes \textit{FALSE}, the loop will be terminated because of the \textit{AND} statement.

\subsection{backward jump}
\textit{Backward jump} is usually used similarly to a while loop and can be easily replaced by it (considering programs that are compiled and executed). Substitution will look like \textbf{do-while} construction.

\begin{figure}[h]
    \centering
    \includegraphics{backward1}
    \caption{\textit{backward jump} scheme}
    \label{fig:my_label}
\end{figure}

@@

\textbf{Goto backward} will work as long as the condition in if-statement is TRUE. In this case we could easily replace construction using \textbf{goto} with the \textit{while}:

\begin{figure}[h]
    \centering
    \includegraphics{backward2}
    \caption{\textit{backward jump} elimination}
    \label{fig:my_label}
\end{figure}

@@

The first iteration will always work, and all subsequent ones only as long as the condition in the loop is \textit{TRUE}.

\subsubsection{continue}
\textit{Continue} is also a control-flow statement, which occurs in loops and is implemented through the EO object \textbf{GOTO}. In order to replace it, we need to invert the \textit{if}-condition in the loop and all the statements that occur after \textbf{goto} are put in the \textit{if} body with the inverted condition.

@@

Since the \textbf{while} loop is an attribute of the \textbf{goto} object, \textit{if-condition} is executed a "jump up" that emulates the \textit{continue} statement. To get rid of this "jump" we should replace the GOTO object with the method described above:
\begin{lstlisting}
[]
 memory > i
 seq > @
 while.
  i.less 10
  seq
   i.write (i.add 1)
   if.
    not. (eq. (i.mod 5 0))
    stdout "Not divisible by 5!\n"
    TRUE
\end{lstlisting}
